Extending the Advanced Forensic Format to accommodate Multiple Data Sources, Logical Evidence, Arbitrary Information and Forensic Workflow

### 2009-03-15 paper/paper.tex
Added a screenshot of windows explorer openning an AFF4 file.

\begin{abstract}
Forensics analysis requires the acquisition and management of many
different types of evidence material, including individual disk
drives, RAID sets, network packets, memory images, and extracted
files. Often the same evidence is reviewed by several different tools
or examiners in different locations. We propose a backwards-compatible
evolutionary redesign of the Advanced Forensic Format an open,
extensible file format for storing and sharing of evidence, arbitrary
case related information and analysis results among different
tools. The new specification was designed to be simple to implement,
allowing the use of the well supported Zip File format specifications
for bit level file access.
\end{abstract}

@@

\subsection{Prior Work}
In recent years there has been a steady and growing interest in the
actual file formats and containers used to store this digital
evidence. Early practitioners created exact bit-for-bit copies
(``mirror disks'') employed proprietary software systems
(e.g. \cite{safeback,ilook,encase}) for making and authenticating
``images'' of digital evidence. PyFlag\cite{pyflag} introduced a
``seekable gzip'' format that allowed disk images to be stored in a
form that was compressed but allowed the random-access to evidence
data necessary for forensic analysis. The \emph{sgzip} format does not
store any metadata at all.

The Advanced Forensic Format (AFF) expanded on this idea with a
forensic file format that allowed both data and arbitrary metadata to
be stored in a single digital archive\cite{garfinkel:aff}. The EWF
file format also provides for a small number of predefined metadata
fields to be stored within the file format.

@@

This paper extends the previous work by adapting AFF to the multiple
heterogeneous data types that might arise in a modern digital
investigation, including data from multiple data storage devices, new
data types (including network packets and memory images), extracted
logical evidence, and forensic workflow. We do this in a manner that
is at once upwards compatible with AFF while adopting the AFF
bit-level specification so that AFF files can be processed using
standard non-forensic tools. The  We call the new system AFF4, and use the
phrase AFF1 to refer to the legacy system developed by
Garfinkel.\footnote{Although Garfinkel never changed the AFF bit-level
specification, Garfinkel released AFFLIB implementations with major
version numbers 1, 2 and 3. We therefore call our system AFF4 to avoid
confusion.}

@@

\item Although the AFF1 specification calls for a ``table of contents'' similar
  to the Zip\cite{zip-format} ``central directory'' that is stored at
  the end of AFF files, Garfinkel never implemented this directory in
  the publicly released AFF1 implementation, AFFLIB. As a result,
  every header of every segment in an AFF file needs to be read when a
  file is opened. It practice this can take 10--30 seconds the first
  time a large AFF file is opened. (Once the file is opened, the
  sectors corresponding to the segment headers are stored in the
  computer's disk cache.)

\item AFF1's bit-level specification is essentially a simple container
  file specification. Given the that there are other container file
  specifications that are much more widely supported with both
  developer and end-user tools, it seemed reasonable to migrate AFF
  from its home-grown format to one of the existing standards.

@@

Consider the typical usage scenario depicted in Figure \ref{usage}, of
an file which contains a disk image. This volume is distributed to two
independent analysts, Alice and Bob. Alice may find and extract
individual files, while Bob may correlate information in the evidence
file with other data that is available on departmental
servers. Although in some environments Alice and Bob may be able to
work on a shared file that is located on a server, in other
environments there will not be sufficient bandwidth or
connectivity. Instead, each analyst will be required to store the
information in their own evidence file; these files will then be
recombined at a later point in time.

@@

\begin{figure}[tb]
  \begin{center}
  \mbox{\epsfxsize=0.6\columnwidth \epsffile{usage1.eps}}
  \caption{A typical usage scenario. Both Alice and Bob receive an AFF
  volume but work independently. Rather than modifying the volume,
  they each create their own local volumes and save their results into
  those files. They can now exchange the smaller new volumes and
  effectively merge their results into the same AFF set when they are finished.}
  \label{usage}
  \end{center}
\end{figure}

@@

We replaced AFF's ``bit-level'' format with an improved format that is
more flexible and has better existing support, namely the Zip64 file
format is not adopted as a default storage format. This enhanced our
compatibility with other tools and simplifies implementations for
compatible systems.

@@

We have also adopted AFFLIB to implement the new schema and file
format, which will allow existing AFF-aware programs to access images
written in either the old or the new format without modification. We
are extending the AFFLIB API to support the new information mode.

@@

\begin{itemize}
\item \emph{An AFF Object} is the basic building block of our
file format. AFF Objects have a globally unique name (URN) as described in
\cite{RFC1737}. The name is defined within the aff2 namespace, and is
made unique by use of a unique identifier generated as per
RFC4122\cite{RFC4122}.

@@

\item \emph{A segment} is a single unit of data written to a volume. AFF4
  segments have a \emph{segment name} provided by their URN, a
  \emph{segment timestamp} in GMT, and the \emph{segment contents}.

@@

\item \emph{The Resolver} is a central data store which collects and resolves
attributes for the different AFF Objects. The Resolver has universal
visibility of objects from all volumes, and therefore guides
implementations in resolving external references. We discuss the
Resolver in detail in Section \ref{resolver}.
\end{itemize}

\section{Metadata and the Universal Resolver}
\label{resolver}
One of the fundamental tasks of forensics is in ensuring the
provenance of any particular piece of digital evidence, for example by
employing a chain of custody process. To assist in this, acquisition
methodologies emphasise the process of individuating evidence, and
storing it in a manner that it is easily identifiable.

Management of evidence requires on effective identififaction, with
practitioners currently employing acquisition time metadata such as
case identifiers and description fields in the EWF file format; file
and directory naming schemes, and labeling of evidence container hard
drives. Evidence may also be referred to by external means in an
inconsistent way. For example, in an investigator's case note a disk
image may be referred to by the name of the suspect (e.g. Joe's hard
disk), the case number or dates.

Such individuation schemes may be, however, problematic when
automatically managing evidence. For example, at acquisition time a
suitably unique individuator may not be selected. If that occured, at
analysis time evidence container files may need to be renamed to avoid
name collisions.

@@

In this way we can identify uniquely any evidence without the change
for name collisions. Once a universal identification regime is
created, it is now possible to collect metadata about each object. The
AFF4 model treats metadata as an abstract concept which may exist
independently from the data itself. We term {\em metadata} to be a set
of statements about objects, written in triple notations (Subject,
Attribute, Value), where {\em Subject} is the URN of the object the
statement is made about. Using this system we are able to store
arbitrary attributes about any object in the AFF4 universe. Additionally,
as these statements are universally scoped, they may be stored anywhere.

@@

\begin{itemize}
\item There is already wide support for the Zip and Zip64 formats. By
  migrating to these formats, we can take advantage of the rich number
  of user and developer tools already available. The volume may be
  inspected using any number of commercial zip application
  (e.g. Windows Explorer natively supports Zip files).

@@

\begin{figure}[tbp]
  \begin{center}
  \mbox{\epsfxsize=0.8\columnwidth \epsffile{zip_structure.eps}}
  \caption{The basic structure of a Zip archive. Also can be seen how
new archive members are added to an existing Zip File. The
Central Directory gets overwritten by the new member, and a new
Central Directory is written on the end.  }
  \label{zip_structure}
  \end{center}
\end{figure}

@@

@@

Similarly, each component can be compressed independently into its own
stream. A single map stream can then be produced to combine all the
component streams into a single logical stream. This approach can take
advantage of multiple systems to actually do the compression as each
component is compressed independently.

@@

\subsection{Forensic Application State}
Often forensic applications need to store files other than the
evidence itself - for example, they might need to store internal data,
annotations, keyword hits etc. Currently these applications store the
data in a proprietary external file or database. This makes it
difficult to archive because the evidence itself may become separated
from the case file.

It is advantageous for these applications to store their state within
the evidence file itself using a Private Stream. Then when opening the
evidence file again, the results of their analysis will become
available.

@@

>>>>

\begin{abstract}
Forensic analysis requires the acquisition and management of many
different types of evidence, including individual disk drives, RAID
sets, network packets, memory images, and extracted files. Often the
same evidence is reviewed by several different tools or examiners in
different locations. We propose a backwards-compatible evolutionary
redesign of the Advanced Forensic Format---an open, extensible file
format for storing and sharing of evidence, arbitrary case related
information and analysis results among different tools. The new
specification was designed to be simple to implement, allowing the use
of the well supported Zip File format specifications for bit level
file access.
\end{abstract}

@@

\subsection{Prior Work}
In recent years there has been a steady and growing interest in the
actual file formats and containers used to store this digital
evidence. Early practitioners created exact bit-for-bit copies
(``mirror disks'') or employed proprietary software systems
(e.g. \cite{safeback,ilook,encase}) for making and authenticating
``images'' of digital evidence. PyFlag\cite{pyflag} introduced a
``seekable gzip'' format that allowed disk images to be stored in a
form that was compressed but allowed random-access to evidence data
necessary for forensic analysis. The \emph{sgzip} format does not
store any metadata at all.

The Advanced Forensic Format (AFF) expanded on this idea with a
forensic file format that allowed both data and arbitrary metadata to
be stored in a single digital archive\cite{garfinkel:aff}. The EWF
file format also provides for a small number of predefined metadata
fields to be stored within the file format\cite{libewf}.

@@

This paper extends the previous work by adapting AFF to the multiple
heterogeneous data types that might arise in a modern digital
investigation, including data from multiple data storage devices, new
data types (including network packets and memory images), extracted
logical evidence, and forensic workflow. We do this in a manner that
is at once upwards compatible with AFF while adopting the AFF
bit-level specification so that AFF files can be processed using
standard non-forensic tools. We call the new system AFF4, and use the
phrase AFF1 to refer to the legacy system developed by
Garfinkel.\footnote{Although Garfinkel never changed the AFF bit-level
specification, Garfinkel released AFFLIB implementations with major
version numbers 1, 2 and 3. We therefore call our system AFF4 to avoid
confusion.}

@@

\item Although the AFF1 specification calls for a ``table of contents'' similar
  to the Zip\cite{zip-format} ``central directory'' that is stored at
  the end of AFF files, Garfinkel never implemented this directory in
  the publicly released AFF1 implementation, AFFLIB. As a result,
  every header of every segment in an AFF file needs to be read when a
  file is opened. In practice this can take 10--30 seconds the first
  time a large AFF file is opened. (Once the file is opened, the
  sectors corresponding to the segment headers are stored in the
  computer's disk cache.)

\item AFF1's bit-level specification is essentially a simple container
  file specification. Given that there are other container file
  specifications that are much more widely supported with both
  developer and end-user tools, it seemed reasonable to migrate AFF
  from its home-grown format to one of the existing standards.

@@

Consider the typical usage scenario depicted in Figure \ref{usage}, of
a volume containing a disk image. This volume is distributed to two
independent analysts, Alice and Bob. Alice may find and extract
individual files, while Bob may correlate information in the evidence
file with other data that is available on departmental
servers. Although in some environments Alice and Bob may be able to
work on a shared file that is located on a server, in other
environments there will not be sufficient bandwidth or
connectivity. Instead, each analyst will be required to store the
information in their own evidence file; these files will then be
recombined at a later point in time.

@@

\begin{figure}[tb]
  \begin{center}
  \mbox{\epsfxsize=0.9\columnwidth \epsffile{usage1.eps}}
  \caption{A typical usage scenario. Both Alice and Bob receive an AFF
  volume but work independently. Rather than modifying the volume,
  they each create their own local volumes and save their results into
  those files. They can now exchange the smaller new volumes and
  effectively merge their results into the same AFF set when they are finished.}
  \label{usage}
  \end{center}
\end{figure}

@@

We replaced AFF's ``bit-level'' format with an improved format that is
more flexible and has better existing support, namely the Zip64 file
format is now adopted as a default storage format. This enhanced our
compatibility with other tools and simplifies implementations for
compatible systems.

@@

We have also adopted AFFLIB to implement the new schema and file
format, which will allow existing AFF-aware programs to access images
written in either the old or the new format without modification. We
are extending the AFFLIB API to support the new information model.

@@

\begin{itemize}
\item \emph{An AFF Object} is the basic building block of our
file format. AFF Objects have a globally unique name (URN) as described in
\cite{RFC1737}. The name is defined within the aff4 namespace, and is
made unique by use of a unique identifier generated as per
RFC4122\cite{RFC4122}.

@@

\label{definitions}
\item \emph{A segment} is a single unit of data written to a volume. AFF4
  segments have a \emph{segment name} provided by their URN, a
  \emph{segment timestamp} in GMT, and the \emph{segment
contents}. Segments are suitable for storing small quantities of data,
and still present a stream interface.

@@

\item \emph{The Resolver} is a central data store which collects and resolves
attributes for the different AFF Objects. The Resolver has universal
visibility of objects from all volumes, and therefore guides
implementations in resolving external references. 
\end{itemize}

\section{Metadata and the Universal Resolver}
\label{resolver}
One of the fundamental tasks of forensics is in ensuring the
provenance of any particular piece of digital evidence, for example by
employing a chain of custody process. To assist in this, acquisition
methodologies emphasise the process of individuating evidence, and
storing it in a manner that is easily identifiable.

Management of evidence requires an effective identififaction, with
practitioners currently employing acquisition time metadata such as
case identifiers and description fields in the EWF file format; file
and directory naming schemes, and labeling of evidence container hard
drives. Evidence may also be referenced by external means in an
inconsistent way. For example, in an investigator's case note a disk
image may be referred to by the name of the suspect (e.g. Joe's hard
disk), the case number or dates.

Such individuation schemes may be problematic when automatically
managing evidence. For example, at acquisition time a suitably unique
individuator may not be selected. If that occured, at analysis time
evidence container files may need to be renamed to avoid name
collisions.

@@

In this way we can identify uniquely any evidence without the chance
for name collisions. Once a universal identification regime is
created, it is now possible to collect metadata about each object. The
AFF4 model treats metadata as an abstract concept which may exist
independently from the data itself. We term {\em metadata} to be a set
of statements about objects, written in triple notations (Subject,
Attribute, Value), where {\em Subject} is the URN of the object the
statement is made about. Using this system we are able to store
arbitrary attributes about any object in the AFF4 universe. Additionally,
as these statements are universally scoped, they may be stored anywhere.

@@

\begin{itemize}
\item There is already wide support for the Zip and Zip64 formats. By
  migrating to these formats, we can take advantage of the rich number
  of user and developer tools already available. The volume may be
  inspected using any number of commercial zip application
  (e.g. Windows Explorer natively supports Zip files as can be seen in
  Figure~\ref{explorer}).

@@

\begin{figure}[tbp]
  \begin{center}
  \mbox{\epsfxsize=0.8\columnwidth \epsffile{zip_structure.eps}}

  \caption{The basic structure of a Zip archive. Also shows is how new
  archive members are added to an existing Zip File. The Central
  Directory gets overwritten by the new member, and a new Central
  Directory is written on the end.  }

  \label{zip_structure}
  \end{center}
\end{figure}

@@

\begin{figure}[tbp]
  \begin{center}
  \mbox{\epsfxsize=0.8\columnwidth \epsffile{compressed_folders.eps}}

  \caption{An Image stream browsed from Windows Explorer.  Basic
  access to the evidence volume can be made using familiar tools
  improving transparency.}

  \label{explorer}
  \end{center}
\end{figure}

@@

Similarly, each component can be compressed independently into its own
stream. A single map stream can then be produced to combine all the
component streams into a single logical stream. This approach can take
advantage of multiple systems to actually do the compression in
parallel as each component is compressed independently.

@@

@@

\subsection{Logical File Acquisition}
Sometimes in a forensic investigation it is impossible to take a true
forensic image of the underlying disk, and we must resort to acquiring
discrete files instead. Such a case might occur for example, when
investigating an incident on a very large network storage system or a
critical system which can not be brought down for the duration of the
image acquisition.

It is still advantageous in this case to bring these files into the
AFF4 evidence universe by acquiring each evidence file using a unique
URN. As explained in \ref{definitions}, segments are AFF4 stream
objects which are implemented by storing the data in Zip archive
members. It follows, therefore, that a regular zip file containing
files is also a valid AFF4 volume providing the segment names are
globally unique.

So a logical image of files, can be created by any regular Zip
compression program in the field. Once brought into the lab these
volumes can be given a volume URN and imported into the Universal
Resolver to provide access to all the files within the archive.
